---
title: "08-EDA of fullscreen results"
author: "Andrew Habrich"
date: "`r Sys.Date()`"
output: html_document
---
```{css zoom-clickfig-src, echo = FALSE}
script src = "https://ajax.googleapis.com/ajax/libs/jquery/3.4.1/jquery.min.js"
```

```{js zoom-clickfig-jquery, echo = FALSE}
 $(document).ready(function() {
    $('body').prepend('<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>');
    // onClick function for all plots (img's)
    $('img:not(.zoomImg)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src')).css({width: '100%'});
      $('.zoomDiv').css({opacity: '1', width: 'auto', border: '1px solid white', borderRadius: '5px', position: 'fixed', top: '50%', left: '50%', marginRight: '-50%', transform: 'translate(-50%, -50%)', boxShadow: '0px 0px 50px #888888', zIndex: '50', overflow: 'auto', maxHeight: '100%'});
    });
    // onClick function for zoomImg
    $('img.zoomImg').click(function() {
      $('.zoomDiv').css({opacity: '0', width: '0%'}); 
    });
  });
```

```{r Scrollable text output, include=FALSE}
options(width = 60)
local({
  hook_output <- knitr::knit_hooks$get('output')
  knitr::knit_hooks$set(output = function(x, options) {
    if (!is.null(options$max.height)) options$attr.output <- c(
      options$attr.output,
      sprintf('style="max-height: %s;"', options$max.height)
    )
    hook_output(x, options)
  })
})
```
<!-- ## Notes --------------------------- -->

<!-- ## > THIS IS JUST THE 'YES' PILE FROM THE TiAb screening phase. -->

<!-- ## > AS OF JULY 13th ALL THE YES HAVE BEEN SCREENED  -->

<!-- ## 1. Load relevant packages-------- -->

## 1. Load packages

```{r packages, message=FALSE, warning=FALSE}
library(tidyverse) #v2.0.0
library(rphylopic) #v1.1.1
library(patchwork) #v1.1.2; arrange figures using patchwork R package ggplot extension
library(maps) #v3.4.1
library(sf) #v1.0-14
library(kableExtra)
```

## 2. Explore data structure

<!-- # rm(list=ls()) #uncomment to remove everything from environment -->

```{r include=FALSE}
fs_yes <- read_csv("./raw_data/full_screening/fullscreen_yes_results.csv")
fs_yes <- fs_yes %>% mutate(ex_r = ifelse(excl_reason=="na", "Keep", "Reject")) #this is a useful category, it can be removed from each of the figure calls and just called directly
str(fs_yes)
```

### 2.1 Criteria DECISIONS

#### 2.1.1 Proportion of keep/reject

```{r}
fs_yes %>% mutate(ex_r = ifelse(excl_reason=="na", "Keep", "Reject")) %>% group_by(ex_r) %>% tally() %>% 
  ggplot(aes(x="", y=n, fill=ex_r)) + 
  geom_bar(width=1, stat= "identity") +
  geom_text(aes(label = n), position = position_stack(vjust = 0.5)) +
  labs(x = "", y = "Count", title = "Fullscreening decision", fill = "") + 
  theme_minimal()
```

#### 2.1.2 By year
```{r}
### create table with count by year of each ex_r group on a stacked bar plot
fs_yes %>% mutate(ex_r = ifelse(excl_reason=="na", "Keep", "Reject")) %>% group_by(pub_year, ex_r) %>% tally() %>% 
  ggplot(aes(x=pub_year, y=n, fill= ex_r)) + 
  geom_bar(width=1, stat= "identity", colour="black") +
  geom_text(aes(label = n), position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values = c("lightgreen","coral"))+ #manually change colours of the bars
  labs(x = "Publication year", y = "Count", title = "Fullscreening decision", fill = "") + 
  theme_minimal() + theme(legend.position = c(0.2,0.8))
```

#### 2.1.3 For the rejections, what criteria did they violate?
```{r include=FALSE}
exl_crit <- fs_yes %>% filter(excl_reason != "na") %>% group_by(excl_reason) %>% count() %>% ungroup() %>% 
  arrange(as.numeric(excl_reason)) %>% mutate(ID = as.integer(excl_reason)) #create an ID column to plot
```

```{r}
#NOTE: See the inclusion criteria list for the description of each of the numbered criteria
exl_crit %>%  
  ggplot(aes(x=ID, y=n)) +
  geom_bar(width=1, stat= "identity", fill = "lightgray", colour= "black") +
  geom_text(aes(label = n), position = position_stack(vjust = 0.8)) +
  labs(x = "Criteria", y = "Count", title = "Rejected entries by criteria") + 
  scale_x_continuous(breaks=seq(1,12,1)) + 
  scale_y_continuous(breaks=seq(0,70,10), expand = c(0, 0), limits = c(0,70)) +
  theme_minimal() + theme(axis.line = element_line(linetype = "solid"),
                          panel.background = element_rect(fill = NA))
```

#### Description of inclusion criteria
```{r}
#create dataframe
incl_criteria <- data.frame(
  Criteria = 1:12,
  Description = c("PDF is inaccessible, through library or other online extraction", 
                  "Document is not in english/Unreliable translation",
                  "No part of the study is in an 'urban area'",
                  "Only PART of the study is in an 'urban area', not the main focus of the study",
                  "No landscape configurational metric (connectivity) is used as an explanatory variable in the study",
                  "The response variable is not related to diversity, abundance, movement, or health of biological organisms",
                  "The focal taxa is not terrestrial for at least a part of its life cycle
",
                  "Biological information is not used to validate the metric of connectivity (purely a description of the landscape) ",
                  "No 'case study' or example given, if the paper was a framework/review/methodology/simulation",
                  "Study reports no relevant extractable information for our study",
                  "Unclear or missing data; Authors need to be contacted",
                  "Review paper, look at the source materials to extract relevant information/articles")
)

# Convert data frame to table using kable()
knitr::kable(incl_criteria, 
             col.names = c("Criteria","Description"),
             align = "cl",
             caption = "Description of inclusion criteria used for full-screening articles") %>% 
  kable_styling("striped")
```

### 2.2 Authors  ####
```{r include=FALSE}
str(fs_yes)
## create a count of the # of authors on each paper
fs_yes$num_authors <- sapply(strsplit(fs_yes$author, ";\\s*|\\s+and\\s+"), length)
```

#### 2.2.1 plot \# of authors by year.

<!-- # Is there a trend of increasing number of authors? -->

<!-- # How many are KEEP/REJECT -->

```{r}
fs_yes %>% mutate(ex_r = ifelse(excl_reason=="na", "Keep", "Reject")) %>% group_by(pub_year) %>% 
  ggplot(aes(x=as.factor(pub_year), y=num_authors, fill = ex_r)) + 
  geom_boxplot(position="dodge", alpha=0.5) +
  labs(x = "Publication year", y = "Avg. Num. Authors", fill="") + 
  scale_fill_manual(values = c("lightgreen","coral")) +
  theme_minimal() + theme(legend.position = c(0.2,0.8),
                        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=0))
```

### 2.3 Journals  ####

```{r}
fs_yes %>% mutate(ex_r = ifelse(excl_reason=="na", "Keep", "Reject")) %>% group_by(journal, ex_r) %>% tally() %>% 
  #summarize(mean_n_auth = mean(num_authors)) %>% ungroup() %>% 
  ggplot(aes(x = n, y = reorder(journal, n), fill = ex_r)) + #reorder journals by descending count
  geom_bar(stat= "identity", colour="black") +
  labs(x = "Count", y = "Journal", fill="") + 
  scale_y_discrete(labels = function(x) str_trunc(x, 35))+
  scale_fill_manual(values = c("lightgreen","coral")) + #manually change colours of the bars
  theme_minimal() +
  theme(axis.text.y = element_text(size = 5),
        legend.position = c(0.8,0.8)) #specific modifications to the theme
```

### 2.4 Taxonomic group

```{r include=FALSE}

## These are not all defined to the same level, so some broader grouping may be required
### CREATE higher level group to bin them into
tax <- fs_yes %>%
  mutate(group = case_when( #when taxa is %in% the list it will create a new column with the higher level grouping
    taxa %in% c("amphibia", "anuran") ~ "amphibian",
    taxa %in% c("reptilia") ~ "reptilian",
    taxa %in% c("aves") ~ "aves",
    taxa %in% c("aranaea", "arthropod","coleoptera", "collembola", "formicidae", "hymenoptera", "lepidoptera") ~ "arthropoda",
    taxa %in% c("chiroptera", "mammalia", "marsupialia", "primate", "rodentia", "scuiridae") ~ "mammalia",
    taxa %in% c("bacteria", "fungi", "gastropoda", "invertebrates") ~ "other",
    taxa %in% c("plants", "pteridophyta") ~ "plants",
    taxa %in% c("mammalia, reptilia", "multi-taxa", "amphibia, reptilia") ~ "multi-taxa",
    TRUE ~ "na")) 

```

#### 2.4.1 how many total papers on each of the taxa groups?

```{r}
tax %>% group_by(group) %>% tally() %>% 
  ggplot(aes(x=group, y=n, fill=group)) +
  geom_bar(width=1, stat= "identity", colour= "black") +
  geom_text(aes(label = n), position = position_stack(vjust = 0.5)) +
  labs(x = "", y = "Count", title = "Number of papers by taxonomic focus", fill = "") + 
  theme_minimal() + theme(legend.position = "none",
                          axis.text.x = element_text(angle = 45, vjust = 1, hjust=0.75))
```

#### 2.4.2 how many total per taxa, for the ones to keep?

```{r}
tax %>% filter(excl_reason == "na") %>% group_by(group) %>% tally() %>% 
  ggplot(aes(x=group, y=n, fill=group)) + 
  geom_bar(width=1, stat= "identity", colour= "black") +
  geom_text(aes(label = n), position = position_stack(vjust = 0.5)) +
  labs(x = "", y = "Count", title = "Taxonomic focus of INCLUDED papers", fill = "") + 
  theme_minimal()+ theme(legend.position = "none",
                         axis.text.x = element_text(angle = 45, vjust = 1, hjust=0.75))
```

#### 2.4.3 stacked figure

```{r}
tax %>% mutate(ex_r = ifelse(excl_reason=="na", "Keep", "Reject")) %>% group_by(group, ex_r) %>% tally() %>% 
  ggplot(aes(x=group, y=n, fill= ex_r)) + 
  geom_bar(width=1, stat= "identity", colour="black") +
  geom_text(aes(label = n), position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values = c("lightgreen","coral"))+ #manually change colours of the bars
  labs(x = "Taxon grouping", y = "Count", title = "Fullscreening decision by taxa", fill = "") + 
  theme_minimal() + theme(legend.position = c(0.8,0.8), #(x,y coords)
                          axis.text.x = element_text(angle = 45, vjust = 1, hjust=0.75)) +
    add_phylopic(name = "Leiopelma archeyi", x = 1, y = 30, ysize = 5)+
    add_phylopic(name = "Danaus plexippus", x = 2, y = 75, ysize = 6)+
    add_phylopic(name = "songbirds", x = 3, y = 95, ysize = 5)+ 
    add_phylopic(name = "Tamiasciurus hudsonicus", x = 4, y = 65, ysize = 5)+ 
    add_phylopic(name = "Tropidodiscus", x = 7, y = 10, ysize = 4)+
    add_phylopic(name = "Acer rubrum", x = 8, y = 55, ysize = 6)+
    add_phylopic(name = "Ardeosaurus brevipes", x = 9, y = 15, ysize = 6)
```

### 2.5 Direction of effect 
#### 2.5.1 ALL TAXA

```{r}
tax %>% filter(excl_reason == "na") %>% group_by(result_sign) %>% tally() %>% 
  ggplot(aes(x = n, y = reorder(result_sign, n), fill=result_sign)) + #reorder journals by descending count
  geom_bar(stat= "identity", colour="black") +
  geom_text(aes(label = n), position = position_stack(vjust = 0.5)) +
  labs(x = "Count", y = "Connectivity metric type used") + 
  scale_fill_manual(values = c(`0` = "#FFFFFF", mixed = "#93AA00", na = "#00C19F", neg = "#F45F5F", pos = "#83FF1B"))+
  scale_y_discrete(labels = function(x) str_trunc(x, 35))+
  theme_minimal() +
  theme(axis.text.y = element_text(size = 11),
        legend.position = "none") +
  coord_cartesian(xlim = c(0, 70)) +
  add_phylopic(uuid="acf1cbec-f6ef-4a82-8ab5-be2f963b93f5", x = 60, y = 1, ysize = 0.5) 
```

#### 2.5.2 TAXA FACETS

```{r}
tax %>% filter(ex_r == "Keep") %>% group_by(group, result_sign) %>% tally() %>% 
  ggplot(aes(x = result_sign, y=n, fill = result_sign)) +
  geom_bar(stat = "identity",colour = "black") +
  geom_text(aes(label = n), position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values = c(`0` = "#FFFFFF", mixed = "#93AA00", na = "#00C19F", neg = "#F45F5F", pos = "#83FF1B")) +
  theme_minimal() + theme(legend.position = "none") +
  facet_wrap(~group) 
```

#### 2.5.3 Make a figure for each group

```{r include=FALSE}
ttal <- tax %>% filter(ex_r == "Keep") %>% group_by(group, result_sign) %>% tally() %>% ungroup() %>% 
  add_row(group = "other", result_sign = "0", n = 0)  %>% 
  add_row(group = "other", result_sign = "neg", n = 0)  %>% 
  add_row(group = "reptilian", result_sign = "0", n = 0) %>%  
  add_row(group = "multi-taxa", result_sign = "0", n = 0)

av <- ttal %>% filter(group == "aves") %>% 
  ggplot(aes(x = result_sign, y=n, fill = result_sign)) + geom_bar(stat = "identity",colour = "black") + 
  add_phylopic(name = "songbirds", x = 1, y = 20, ysize = 4)
ma <- ttal %>% filter(group == "mammalia") %>% 
  ggplot(aes(x = result_sign, y=n, fill = result_sign)) + geom_bar(stat = "identity",colour = "black") + 
  add_phylopic(name = "Tamiasciurus hudsonicus", x = 1, y = 20, ysize = 4)

ar <- ttal %>% filter(group == "arthropoda") %>% 
  ggplot(aes(x = result_sign, y=n, fill = result_sign)) + geom_bar(stat = "identity",colour = "black") + 
  add_phylopic(name = "Danaus plexippus", x = 1, y = 25, ysize = 5)
pl <- ttal %>% filter(group == "plants") %>% 
  ggplot(aes(x = result_sign, y=n, fill = result_sign)) + geom_bar(stat = "identity",colour = "black") + 
  add_phylopic(name = "Acer rubrum", x = 1, y = 25, ysize = 4)

re <- ttal %>% filter(group == "reptilian") %>% 
  ggplot(aes(x = result_sign, y=n, fill = result_sign)) + geom_bar(stat = "identity",colour = "black") + 
  add_phylopic(name = "Ardeosaurus brevipes", x = 1, y = 10, ysize = 3)
am <- ttal %>% filter(group == "amphibian") %>% 
  ggplot(aes(x = result_sign, y=n, fill = result_sign)) + geom_bar(stat = "identity",colour = "black") + 
  add_phylopic(name = "Leiopelma archeyi", x = 1, y = 10, ysize = 2)

mt <- ttal %>% filter(group == "multi-taxa") %>% 
  ggplot(aes(x = result_sign, y=n, fill = result_sign)) + geom_bar(stat = "identity",colour = "black") + 
  add_phylopic(name = "Homo sapiens", x = 1, y = 10, ysize = 4)
ot <- ttal %>% filter(group == "other") %>% 
  ggplot(aes(x = result_sign, y=n, fill = result_sign)) + geom_bar(stat = "identity",colour = "black") + 
  add_phylopic(name = "Tropidodiscus", x = 1, y = 10, ysize = 2)
```

```{r echo=FALSE, fig.height=11, fig.width=7}
## arrange them all together in one plot (Use & or * to apply theme settings to ALL the plots)
(av | ma ) /
(ar | pl ) /
(re | am ) /
(mt | ot)   & 
geom_text(aes(label = n), position = position_stack(vjust = 0.5)) &
scale_fill_manual(values = c(`0` = "#FFFFFF", mixed = "#93AA00", na = "#00C19F", neg = "#F45F5F", pos = "#83FF1B")) &
theme_minimal() & theme(legend.position = "none") & labs(x="")
```

### 2.6 Other data
#### 2.6.1 validation data
```{r}
##  PROBLEMS WITH SOME ENTRIES; NEEDS COSOLIDATING IN CSV FILES.
tax %>% filter(excl_reason == "na") %>% group_by(validation) %>% tally() %>% 
  ggplot(aes(x = n, y = reorder(validation, n))) + #reorder journals by descending count
  geom_bar(stat= "identity", colour="black") +
  labs(x = "Count", y = "validation type") + 
  scale_y_discrete(labels = function(x) str_trunc(x, 35))+
  theme_minimal() +
  theme(axis.text.y = element_text(size = 11))
```

#### 2.6.2 connectivity type

```{r}
tax %>% filter(excl_reason == "na") %>% group_by(conn_type) %>% tally() %>% 
  ggplot(aes(x = n, y = reorder(conn_type, n))) + #reorder journals by descending count
  geom_bar(stat= "identity", colour="black") +
  labs(x = "Count", y="", title = "Connectivity metric type used") + 
  scale_y_discrete(labels = function(x) str_trunc(x, 35))+
  theme_minimal() +
  theme(axis.text.y = element_text(size = 11))
```

### 2.7 Country & City

```{r include=FALSE}
## 2.7.0 Prepare Map visualization data
globe <- map_data("world", boundary=T) ##the country names have to match the dataframe for it to plot!
cities<- world.cities #get data on cities from 'maps' package
## create a basemap
basemap <- ggplot() +
  geom_polygon(data = globe, aes(x=long, y=lat, group=group), colour="grey50", fill="grey", alpha=0.3) + 
  theme_minimal() + coord_sf()
```

```{r include=FALSE}
#### 2.7.1 COUNTRY ####
ctal <- tax %>% separate_rows(country, sep = ", ") %>% group_by(country, ex_r) %>% count(country) %>% 
  ungroup(ex_r) %>% mutate(total= sum(n)) #create a total column to arrange by
```

#### 2.7.1-1 Frequency of keep/reject by country

```{r}
ctal %>%  ggplot(aes(y = reorder(country, total), x = n, fill = ex_r)) +
    geom_bar(stat = "identity") +
  scale_fill_manual(values = c(Keep = "lightgreen", Reject = "coral")) +
  labs(x = "Count", y = "Country", fill = "") +
  theme_minimal() + theme(legend.position = c(0.8,0.8))
```

#### 2.7.1-2 just the 'keep's

```{r}
ctal %>% filter(ex_r == "Keep") %>% ggplot(aes(y = reorder(country, n), x = n, fill = ex_r)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c(Keep = "lightgreen")) +
  labs(x = "Count", y = "Country", fill = "") +
  theme_minimal() + theme(legend.position = "none")
```

#### 2.7.1-3 add plot of proportion retained and count by country

```{r}
## merge dataframes with frequencies and coordinates of capital cities
cfreq <- merge(ctal, cities %>% filter(capital == 1), by.x = "country", by.y = "country.etc", all.x = T)
cfreq <- cfreq %>% filter(ex_r=="Keep") %>% mutate(prop = n/total)

basemap +  
  geom_point(data=cfreq, aes(x=long,y=lat, size = total, colour=prop), alpha=0.7) +
  theme(legend.position= "bottom") + labs(x="",y="", size="Count", colour="Prop. retained") +
  scale_color_viridis_c(option = "inferno", direction = 1)
```

#### 2.7.2 CITIES

```{r include=FALSE}
# seperate out the multi-city entries, and count occurrences by city/country
cityfreq <- tax %>% separate_rows(urban_area, sep = "; ") %>% 
                group_by(urban_area, country, ex_r) %>% count(urban_area) %>% 
                ungroup(ex_r) %>% mutate(total= sum(n)) #create a total column to arrange by
# merge with world cities data
cfq <- merge(cityfreq, cities, by.x = "urban_area", by.y = "name", all.x = T)
cfq <- cfq %>% filter(ex_r=="Keep") %>% mutate(prop = n/total)

```

```{r}
#Plot together ###NOTE: Some cities do not have coordinates
basemap +  
  geom_point(data=cfq, aes(x=long,y=lat, size = total, colour=prop), alpha=0.7) +
  theme(legend.position= "bottom") + labs(x="",y="", size="Count", colour="Prop. retained")+
  scale_color_viridis_c(option = "inferno", direction = 1)
```
